---
title: "K23_Aim2_LPV_Culture_CohortClean"
author: "Chad H. Hochberg"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(here) #Makes working with directories easier
library(tidyverse) #dplyr syntax for data wrangling
library(data.table) #Some commands here are faster than dplyr when working with big data
library(arrow) #This library allows working with apache arrow which is a memory efficient way to work with these data
library(collapse) #Fast Data Manipulation Package
library(DescTools) #Help for Descriptive Statistics

#Set Working Directory
knitr::opts_knit$set(root.dir = "~/workspace/Storage/chochbe1/persistent/K23_Aim1_LPV_Analysis")

#Create Commonly Used Function
`%!in%` = Negate(`%in%`)

#Project Path
project_path <- '~/workspace/Storage/chochbe1/persistent/K23_Aim1_LPV_Analysis'

#Here
here::i_am('K23_aim1_lpv_here.R')
```


```{r Open Clinical Dataset and Survey Datasets}
lpv_clinical_data <- read_parquet(paste0(project_path, '/data/lpv_clinical_data.parquet'))
survey_analytic <- fread(paste0(project_path, '/data/survey_analytic.csv'))
survey_unit_level <- fread(paste0(project_path, '/data/survey_unit_level.csv'))
```


#Implement Stage 2 Exclusions for Clinical Data
1. MV Initiated at OSH
2. ECMO within 12 hours of LPV Time Zero
3. Tracheostomy as First Airway and IMV Started on Hospital Day 1
4. Age >= 18 Years
5. Died or Discharged within 12 Hours of LPV Time Zero (should be zero of these based on stage 1 exclusions)

```{r Second stage Exclusions}
cat('\n For Consort: Before Second Stage Exlusions there were', length(unique(lpv_clinical_data$pat_enc_csn_id)), 
    'encounters amongst,',length(unique(lpv_clinical_data$cohort_id)), 'patients')
n_enc <- length(unique(lpv_clinical_data$pat_enc_csn_id))
n_ids <- length(unique(lpv_clinical_data$cohort_id))

#MV at Outside Hospital
lpv_analytic <- lpv_clinical_data |> 
  filter(vent_at_osh==0)

cat('\n For Consort:', n_enc-length(unique(lpv_analytic$pat_enc_csn_id)), 
    'encounters were excluded for MV Initiated at OSH (',n_ids-length(unique(lpv_analytic$cohort_id)), 'patients)')
n_enc <- length(unique(lpv_analytic$pat_enc_csn_id)) #Track for Consort
n_ids <- length(unique(lpv_analytic$cohort_id)) #Track for Consort

#Tracheostomy is First Airway and Present in First 24 Hours of Hospitalization
lpv_analytic <- lpv_analytic |>
  mutate(trach_exclude=fifelse(
    tracheostomy=='Trach First Airway' & first_trach_time<=(final_admit_date+dhours(24)), 1, 0)) |> #Small Number of Airway Variables Missing
  mutate(trach_exclude=fifelse(
    is.na(tracheostomy), 0, trach_exclude
  )) |>
  filter(trach_exclude==0) |>
  select(-trach_exclude)
cat('\n For Consort:', n_enc-length(unique(lpv_analytic$pat_enc_csn_id)), 
    'encounters were excluded for Tracheostomy as First Airway Present on Day 1 (',n_ids-length(unique(lpv_analytic$cohort_id)), 'patients)')
n_enc <- length(unique(lpv_analytic$pat_enc_csn_id)) #Track for Consort
n_ids <- length(unique(lpv_analytic$cohort_id)) #Track for Consort

#ECMO Initiated within 12 Hours of LPV Time Zero
#NOTE: Will Keep the Very Small Group of Patients Who Were Not Intubated on VA_ECMO and then Go On Vent AFTER ECMO Support is OFF (for example, for heart transplant)
lpv_analytic <- lpv_analytic |>
  mutate(time_to_ecmo=as.duration(ecmo_start-lpv_time_zero)) |>
  mutate(ecmo_exclude=fcase(
    time_to_ecmo<=dhours(12) & ecmo_end>study_vent_start, 1,
    default = 0
  )) |>
  filter(ecmo_exclude==0) |>
  select(-ecmo_exclude)

cat('\n For Consort:', n_enc-length(unique(lpv_analytic$pat_enc_csn_id)), 
    'encounters were excluded for VA or VV ECMO started within 12 Hours of Study Vent Start (',n_ids-length(unique(lpv_analytic$cohort_id)),'patients)')
n_enc <- length(unique(lpv_analytic$pat_enc_csn_id)) #Track for Consort
n_ids <- length(unique(lpv_analytic$cohort_id)) #Track for Consort

#No Height Available
lpv_analytic <- lpv_analytic |>
  filter(!is.na(height_inches))
cat('\n For Consort:', n_enc-length(unique(lpv_analytic$pat_enc_csn_id)), 
    'encounters were excluded for no Available Height Measurements (',n_ids-length(unique(lpv_analytic$cohort_id)),'patients)')
n_enc <- length(unique(lpv_analytic$pat_enc_csn_id)) #Track for Consort
n_ids <- length(unique(lpv_analytic$cohort_id)) #Track for Consort

#Discharged Within 12 Hours of LPV Time Zero
#As Expected this is 0
lpv_analytic <- lpv_analytic |>
  mutate(dc_exclude=fcase(
    is.na(final_dc_date) & is.na(study_vent_end), 1,
    as.duration(final_dc_date-lpv_time_zero)<dhours(12), 1,
    default=0
  ))
cat('\n For Consort:', n_enc-length(unique(lpv_analytic$pat_enc_csn_id)), 
    'encounters were excluded for Discharge within 12 Hours of LPV Time Zero (',n_ids-length(unique(lpv_analytic$cohort_id)),'patients)')
n_enc <- length(unique(lpv_analytic$pat_enc_csn_id)) #Track for Consort
n_ids <- length(unique(lpv_analytic$cohort_id)) #Track for Consort

#Age Less Than 18
lpv_analytic <- lpv_analytic |>
  filter(age_at_admission>=18)
cat('\n For Consort:', n_enc-length(unique(lpv_analytic$pat_enc_csn_id)), 
    'encounters were excluded Age < 18 years (',n_ids-length(unique(lpv_analytic$cohort_id)),'patients)')
n_enc <- length(unique(lpv_analytic$pat_enc_csn_id)) #Track for Consort
n_ids <- length(unique(lpv_analytic$cohort_id)) #Track for Consort

#Temporary List of COhort ID and Encounter Block
temp_cohort <- lpv_analytic |>
  select(cohort_id, encounter_block, study_vent_start, study_vent_end, lpv_time_zero) |>
  mutate(in_cohort=1)

```

```{r Process Long Tidal Volume Data Into 72 Hours and 24 Hour Datasets}
vent72 <- open_dataset(paste0(project_path, '/data/ventilator_long_72hours.parquet')) |> compute()

#Filter to Those in the Cohort
vent72_analytic <- vent72 |>
  left_join(temp_cohort) |>
  filter(in_cohort==1) |>
  select(-in_cohort) |>
  compute()
  
rm(vent72, temp_cohort)

vent24_analytic <- vent72_analytic |> collect() |>
  #Create an End Time for LPV Assessment at 24 Hours (the earliest of ventilator end or 24 hours after lpv time zero)
  mutate(time_24hours=lpv_time_zero+dhours(24)) |>
  mutate(lpv_end_24hours=fifelse(
  time_24hours<=study_vent_end, time_24hours, study_vent_end)) |>
  select(-time_24hours) |>
  filter(recorded_time<=lpv_end_24hours)
  
#Implement Final Inclusion Criteria
#No Eligible Tidal Volumes in First 24 Hours
eligible_tv_cohort <- vent24_analytic |>
  group_by(cohort_id, encounter_block) |>
  filter(sum(eligible_tv, na.rm = TRUE)>=1) |>
  filter(row_number()==1) |>
  ungroup() |>
  mutate(yes_tv=1) |>
  select(cohort_id, encounter_block, yes_tv) 

lpv_analytic <- lpv_analytic |>
  left_join(eligible_tv_cohort) |>
  filter(yes_tv==1) |>
  select(-yes_tv)

cat('\n For Consort:', n_enc-length(unique(lpv_analytic$pat_enc_csn_id)), 
    'encounters were excluded for no Eligible TV Measurements in First 24 Hours (',n_ids-length(unique(lpv_analytic$cohort_id)),'patients)')
n_enc <- length(unique(lpv_analytic$pat_enc_csn_id)) #Track for Consort
n_ids <- length(unique(lpv_analytic$cohort_id)) #Track for Consort

#Finally, For Patients with More Than 1 Encounter Randomly Select 1 Encounter
set.seed(322)
temp_cohort <- lpv_analytic |>
  select(cohort_id, encounter_block) |>
  group_by(cohort_id) |>
  slice_sample(n=1) |>
  ungroup() |>
  mutate(keep=1) 

lpv_analytic <- lpv_analytic |>
  left_join(temp_cohort) |>
  filter(keep==1) |>
  select(-keep)
  
cat('\n For Consort:', n_enc-length(unique(lpv_analytic$pat_enc_csn_id)), 
    'encounters were excluded after randomly selecting 1 encounter for each eligible patient yielding (',n_ids-length(unique(lpv_analytic$cohort_id)),'patients ecxluded)')
n_enc <- length(unique(lpv_analytic$pat_enc_csn_id)) #Track for Consort
n_ids <- length(unique(lpv_analytic$cohort_id)) #Track for Consort
  
#Final Cohort Numbers
cat('\n For Consort:', n_enc, 
    'encounters in the final cohort (',n_ids,'patients)')
rm(eligible_tv_cohort)
```

```{r Describe the LPV Metrics Available}
final_cohort_ids <- lpv_analytic |>
  select(cohort_id, encounter_block) |>
  mutate(in_cohort=1)

#Filter the Vent Observation Tables to Those in the lpv_analytic dataset
vent72_analytic <- vent72_analytic |>
  left_join(final_cohort_ids) |>
  filter(in_cohort==1) |>
  select(-in_cohort) |>
  compute()
vent24_analytic <- vent24_analytic |>
  left_join(final_cohort_ids) |>
  filter(in_cohort==1) |>
  select(-in_cohort) |>
  compute()

#Primary: % Eligible Time with TV <= 6.5 cc/kg-IBW and Pplat <=30 in 1st 24 hours
#Need to Calculate Time At LPV and Denominator of Eligible Time
#Additionally Calculate Components (time-weighted VT per kg IBW and time-weighted PPlat)

#Track Numbers of Observations for Description in the Paper
cat("\n In the analytic cohort there were,", dim(vent24_analytic)[1], "observed ventilator settings in the first 24 hours")

vent24_outcomes <- vent24_analytic |>
  #How Many Ventilator Measurements in First 24Hours
  group_by(cohort_id) |>
  mutate(n_vent_observations=n()) |>
  #For LPV Assessment Need to Use Paired Measurements of TV and PBW
  mutate(eligible_lpv=fifelse(
    eligible_tv==1 & (!is.na(pplat_lpv) & !is.na(tv_pbw)), 1, 0
  )) |>
  group_by(cohort_id, eligible_lpv) |>
  mutate(n_lpv_measures=fifelse(
    eligible_lpv==1, n(), NaN
  )) |>
  group_by(cohort_id, eligible_tv) |>
  mutate(n_tv_eligible=fifelse(
    !is.na(tv_lpv) & eligible_tv==1, n(), NaN
  )) |>
  mutate(n_pplat_eligible=fifelse(
    !is.na(pplat_lpv) & eligible_tv==1, n(), NaN
  )) |>
  group_by(cohort_id) |>
  fill(n_lpv_measures, n_tv_eligible, n_pplat_eligible, .direction = 'updown') |>
  ungroup() |>
  mutate(n_tv_eligible=fifelse(
    is.na(n_tv_eligible), 0, n_tv_eligible
  )) |>
  mutate(n_pplat_eligible=fifelse(
    is.na(n_pplat_eligible), 0, n_pplat_eligible
  ))

#For the Patients With no Eligible PPlat Can Use Ppeak 
cat('\n Patients with no Eligible PPlat in 1st 24 hours:\n')
cat(length(unique(vent24_outcomes$cohort_id[vent24_outcomes$n_pplat_eligible==0])), 'patients \n')
cat('\n Patients with no Eligible LPV Measures in 1st 24 hours:\n')
cat(length(unique(vent24_outcomes$cohort_id[is.na(vent24_outcomes$n_lpv_measures)])), 'patients \n')

#Use Ppeak
vent24_outcomes <- vent24_outcomes |>
  #Keep Track of the Pplats in Which We Use PPeak Imputation
  mutate(pplat_imp_ppeak=fcase(
    eligible_tv==1 & is.na(n_lpv_measures), peak_inspiratory_pressure_obs
  )) |>
  #IF Pplat is Missing and It Was Otherwise an Eligible Measure Use the Ppeak Imputation
  mutate(pplat_lpv=fifelse(
    !is.na(pplat_imp_ppeak) & is.na(pplat_lpv), pplat_imp_ppeak, pplat_lpv
  )) 

#1 Patient Who Only Has 1 Pplat but Is Recorded at the Same Time as lpv_time_24 meaning there are 0 minutes of that included, Will Fix Here
vent24_outcomes <- vent24_outcomes |>
  mutate(temp=fifelse(
    cohort_id=='CRI0ucu5zrau7g' & mode_name=='PRVC', peak_inspiratory_pressure_obs, NaN)) |>
  mutate(pplat_lpv=fifelse(!is.na(temp), temp, pplat_lpv)) |>
  select(-temp)

#Redefine Eligible LPV and Recount the Numbers
vent24_outcomes <- vent24_outcomes |>
  mutate(eligible_lpv=fifelse(
    eligible_tv==1 & (!is.na(pplat_lpv) & !is.na(tv_pbw)), 1, 0
  )) |>
  group_by(cohort_id, eligible_lpv) |>
  mutate(n_lpv_measures=fifelse(
    eligible_lpv==1, n(), NaN
  )) |>
  group_by(cohort_id) |>
  fill(n_lpv_measures, .direction = 'updown') |>
  ungroup() |>
  mutate(n_lpv_measures=fifelse(
    is.na(n_lpv_measures), 0, n_lpv_measures
  ))

#After Imputation Counts
cat('\n In Patients with no Eligible PPlat in 1st 24 hours, We Used PPeak and After That Imputation There were:\n')
cat(length(unique(vent24_outcomes$cohort_id[vent24_outcomes$n_lpv_measures>=1])), 'patients with LPV measures for Analysis \n')
cat('\n Patients with no Eligible LPV Measures in 1st 24 hours:\n',length(unique(vent24_outcomes$cohort_id[vent24_outcomes$n_lpv_measures==0])))

#Comparison of 'Imputed PPlat' and 'Recorded Pplat'
cat('\n Recorded PPlat During Eligible Measurements \n')
summary(vent24_outcomes$pplat_lpv[(is.na(vent24_outcomes$pplat_imp_ppeak) & vent24_outcomes$eligible_lpv==1)])
cat('\n Ppeak Substitutin During PPlat During Eligible Measurements \n')
summary(vent24_outcomes$pplat_imp_ppeak)

cat('\n Of ventilator observations,\n ', dim(vent24_outcomes[vent24_outcomes$eligible_lpv == 1, ])[1], 'met criteria for LPV analysis,\n')
cat(' ', dim(vent24_analytic[vent24_analytic$eligible_tv == 1, ])[1], 'met criteria for TV analysis, and \n')
cat(' ', dim(vent24_analytic[(vent24_analytic$eligible_tv == 1 & !is.na(vent24_analytic$pplat_lpv)), ])[1], 'met criteria for Pplat analysis in this study. \n')

cat('\n Summary Statistics for Number of Eligible Measures Per Participant')
cat('\n Tidal Volume Measures Per Participant (1st 24 Hours) \n')
summary(vent24_outcomes$n_tv_eligible)
cat('\n Number of Pplat Measures Per Participant (1st 24 Hours) \n')
summary(vent24_outcomes$n_pplat_eligible)
cat('\n Number of LPV Measures Per Participant (1st 24 Hours) \n')
summary(vent24_outcomes$n_lpv_measures)

cat('\n Patients Included in the Primary LPV Analysis:\n')
cat(length(unique(vent24_outcomes$cohort_id[vent24_outcomes$n_lpv_measures>=1])), 'patients \n')

#Keep Track of Modes While TV Eligible
cat('\n Ventilator Modes During TV Eligible Periods:\n')
PercTable(table(vent24_outcomes$mode_name[vent24_outcomes$eligible_tv==1]), rfrq = '001')

cat('\n Ventilator Modes During Any Period in 1st 24 Hours\n')
PercTable(table(vent24_outcomes$mode_name), rfrq = '001')
```

#Below will Calculate LPV Outcomes.
Primary: 
- % Eligible Time with TV <= 6.5 cc/kg-IBW and Pplat <=30 in 1st 24 hours
Secondary: 
-Restrictive definition: % Eligible Time with TV <= 6.5 cc/kg-IBW and Pplat <=30 in 1st 24 hours AND if pplat>30 than Vt needs to be <6.0 cc/kgIBW
- Hours spent with Vt>=8cc/kg IBW in 1st 24-hours of MV
Descriptive:
-Time-weighted TV, Time-weighted PEEP, TIme-weighted PPlat, Time-weighted Driving Pressure
```{r Further Process LPV Outcomes}
#Calculate Time Spent With Each Measurement
vent24_outcomes <- vent24_outcomes |>
  #Calculate Total TIme Observed by 24 Hours Per Patient
  group_by(cohort_id) |>
  mutate(time_at_measure=fcase(
    row_number()!=n(), as.duration(lead(recorded_time)-recorded_time)/dhours(1), #Time to Next Observation
    row_number()==n(), as.duration(lpv_end_24hours-recorded_time)/dhours(1) #If last observation Time to End of 24 Hour Period
  )) |>
  mutate(total_time_observed24=sum(time_at_measure)) |>
  #Carry Forward PEEP
  mutate(peep_extrapolated=peep_set) |>
  fill(peep_extrapolated, .direction= 'downup') |>
  #Now Total LPV Eligible Time
  mutate(eligible_lpv_time=fcase(
    eligible_lpv==1 | eligible_tv==0, 1, #To calculate time from paired TV/PPlat measurements to next. IF measure not eligible for TV measurement (not on controlled mode) time stops until next measurement is eligible
    eligible_lpv==0 & eligible_tv==1, 2
  )) |>
  group_by(cohort_id, eligible_lpv_time) |>
  mutate(time_at_lpv_measure=fcase(
    row_number()!=n(), as.duration(lead(recorded_time)-recorded_time)/dhours(1), #Time to Next Observation
    row_number()==n(), as.duration(lpv_end_24hours-recorded_time)/dhours(1) #If last observation Time to End of 24 Hour Period
  )) |>
  #Now Sum For Each Patient
  group_by(cohort_id, eligible_lpv) |>
  mutate(time_lpv_eligible24=fcase(
    eligible_lpv_time==1 & eligible_tv==1, sum(time_at_lpv_measure),
    default = NaN
  )) |>
  #Now Can Sum Time at TV
  group_by(cohort_id, eligible_tv) |>
  mutate(time_tv_eligible24=fcase(
    eligible_tv==1, sum(time_at_measure)
  )) |>
  #Total TV Eligible Time (NOTE: For Pplat Can Use Total LPV Time)
  group_by(cohort_id) |>
  fill(time_tv_eligible24, time_lpv_eligible24, .direction='updown') |>
  ungroup()

cat('\n Time (hours) per Patient Available for LPV Analysis [This will be the Same for Driving Pressure and PEEP \n')
summary(vent24_outcomes$time_lpv_eligible24[vent24_outcomes$n_lpv_measures>0])
hist(vent24_outcomes$time_lpv_eligible24[vent24_outcomes$n_lpv_measures>0])

cat('\n Time (hours) per Patient Available for TV Analysis \n')
summary(vent24_outcomes$time_tv_eligible24[vent24_outcomes$n_tv_eligible>0])
hist(vent24_outcomes$time_tv_eligible24[vent24_outcomes$n_tv_eligible>0])

#Pimary Outcome: Percentage of Time At LPV
vent24_outcomes <- vent24_outcomes |>
  mutate(met_lpv=fcase(
    eligible_lpv==1 & tv_pbw<=6.5 & pplat_lpv<=30, 1,
    eligible_lpv==1 & (tv_pbw>6.5 | pplat_lpv>30), 0
  )) |>
  #If patient Didn't Meet LPV THis Won't Be Included in Numerator so Numerator is Time LPV Met and Denominator is Total Eligible Time
  group_by(cohort_id) |>
  mutate(hours_at_lpv=sum((met_lpv*time_at_lpv_measure), na.rm=TRUE)) |>
  mutate(percent_lpv=
           hours_at_lpv/time_lpv_eligible24) |>
  #Repeat With Restricted Definition
  mutate(met_lpv_restrict=fcase(
    eligible_lpv==1 & tv_pbw<=6.5 & pplat_lpv<=30, 1,
    eligible_lpv==1 & tv_pbw<6.0 & pplat_lpv>30, 1,
    eligible_lpv==1 & (tv_pbw>6.5 | (tv_pbw>=6.0 & pplat_lpv>30)), 0
  )) |>
  #If patient Didn't Meet LPV THis Won't Be Included in Numerator so Numerator is Time LPV Met and Denominator is Total Eligible Time
  group_by(cohort_id) |>
  mutate(hours_at_lpv_restrict=sum((met_lpv_restrict*time_at_lpv_measure), na.rm=TRUE)) |>
  mutate(percent_lpv_restrict=
           hours_at_lpv_restrict/time_lpv_eligible24) |>
  ungroup()

#Secondary Outcome Hours (%) of Time Spent with TV >8.0 in 1st 24 Hours, TV > 7 in 1st 24 Hours, TV > 6.5 in 1st 24 Hours
#Now Create Time-weighted TV, PEEP and Driving Pressure Measures
vent24_outcomes <- vent24_outcomes |>
  mutate(tv_gt8=fcase(
    eligible_tv==1 & tv_pbw>8, 1,
    eligible_tv==1 & tv_pbw<=8, 0
  )) |>
  group_by(cohort_id) |>
  mutate(hours_tv_gt8=sum((tv_gt8*time_at_measure), na.rm=TRUE)) |>
  mutate(percent_tv_gt8=
           hours_tv_gt8/time_tv_eligible24) |>
  #Repeat for TV of > 7
  mutate(tv_gt7=fcase(
    eligible_tv==1 & tv_pbw>7, 1,
    eligible_tv==1 & tv_pbw<=7, 0
  )) |>
  group_by(cohort_id) |>
  mutate(hours_tv_gt7=sum((tv_gt7*time_at_measure), na.rm=TRUE)) |>
  mutate(percent_tv_gt7=
           hours_tv_gt7/time_tv_eligible24) |>
  #Repeat for TV of > 6.5
  mutate(tv_gt6.5=fcase(
    eligible_tv==1 & tv_pbw>6.5, 1,
    eligible_tv==1 & tv_pbw<=6.5, 0
  )) |>
  group_by(cohort_id) |>
  mutate(hours_tv_gt6.5=sum((tv_gt6.5*time_at_measure), na.rm=TRUE)) |>
  mutate(percent_tv_gt6.5=
           hours_tv_gt6.5/time_tv_eligible24) |>#Repeat for TV of > 6.5
  #Repeat for TV 6
  mutate(tv_gt6=fcase(
    eligible_tv==1 & tv_pbw>6, 1,
    eligible_tv==1 & tv_pbw<=6, 0
  )) |>
  group_by(cohort_id) |>
  mutate(hours_tv_gt6=sum((tv_gt6*time_at_measure), na.rm=TRUE)) |>
  mutate(percent_tv_gt6=
           hours_tv_gt6/time_tv_eligible24) |>
  ungroup()
    
#Now Time Weighted TV, Pplat, PEEP and Driving Pressure; Baseline Respiratory Compliance
vent24_outcomes <- vent24_outcomes |>
  #Start with Tidal Volume
  group_by(cohort_id) |>
  #Calculate Weighted Tidaly Volume = TV* (time observed/time eligible)
  mutate(temp=fcase(
    eligible_tv==1, tv_pbw*(time_at_measure/time_tv_eligible24)
  )) |>
  #Sum the Non Missing
  mutate(tv_weighted24=sum(temp, na.rm=TRUE)) |>
  mutate(tv_weighted24=round(tv_weighted24, digits = 1)) |>
  #Repeat For Pplat (can use eligible_lpv for this Measure)
  mutate(temp=fcase(
    eligible_lpv==1, pplat_lpv*(time_at_lpv_measure/time_lpv_eligible24)
  )) |>
  #As Along As All Values ARe Not Missing Can Do Sum
  mutate(pplat_weighted24=fcase(
    sum(is.na(temp))!=n(), sum(temp, na.rm=TRUE))) |>
  mutate(pplat_weighted24=round(pplat_weighted24, digits = 0)) |>
  #Repeat For PEEP (can use eligible_tv for this Measure)
  mutate(temp=fcase(
    eligible_tv==1, peep_set*(time_at_measure/time_tv_eligible24)
  )) |>
  mutate(peep_weighted24=fcase(
    sum(is.na(temp))!=n(), sum(temp, na.rm=TRUE))) |>
  mutate(peep_weighted24=round(peep_weighted24, digits = 0)) |>
  #Repeat For Driving Pressure (can use eligible_lpv for this Measure)
  #Filter Out Physiologically Implausible Values (Driving Pressure > 2 allowed)
  mutate(temp_dp=fifelse(
    eligible_lpv==1 & (pplat_lpv-peep_extrapolated>2), 1, 0
  )) |>
  #Filter Out the Time That the Implausible Measures Take
  mutate(temp_time=fcase(
    temp_dp==1, time_at_lpv_measure
  )) |>
  mutate(temp_time=fcase(
    sum(is.na(temp_time))!=n(), sum(temp_time, na.rm = TRUE))) |>
  mutate(temp=fcase(
    temp_dp==1, (pplat_lpv-peep_extrapolated) *(time_at_lpv_measure/temp_time)
  )) |>
  mutate(dp_weighted24=fcase(
    sum(is.na(temp))!=n(), sum(temp, na.rm=TRUE))) |>
  mutate(dp_weighted24=round(dp_weighted24, digits = 0)) |>
  select(-temp_dp, -temp_time) |>
  ungroup()

#For Purposes of Comparison to Other Metrics Get First TV, Mean TV and Median TV
vent24_outcomes <- vent24_outcomes |>
  group_by(cohort_id, eligible_tv) |>
  mutate(tv_first=fcase(
    eligible_tv==1 & row_number()==1, tv_pbw
  )) |>
  mutate(tv_mean24=fcase(
    eligible_tv==1, mean(tv_pbw, na.rm = TRUE)
  )) |>
  mutate(tv_median24=fcase(
    eligible_tv==1, median(tv_pbw, na.rm = TRUE)
  )) |>
  group_by(cohort_id) |>
  fill(tv_first, tv_mean24, tv_median24, .direction = 'updown') |>
  relocate(tv_first, tv_mean24, tv_median24, .after = tv_weighted24)
  
#Calculate Baseline Respiratory Compliance
vent24_outcomes <- vent24_outcomes |>
  #Redefine Eligible Measurements, Will Exclude When PPlat-PEEP<3 as these are outlier not physiologic values
  mutate(temp=fifelse(
    eligible_lpv==1 & (pplat_lpv-peep_extrapolated>2), 1, 0
  )) |>
  group_by(cohort_id, temp) |>
  mutate(resp_compliance_baseline=fcase(
    row_number()==1 & temp==1, tv_lpv/(pplat_lpv-peep_extrapolated)
  )) |>
  mutate(resp_compliance_baseline=round(resp_compliance_baseline, digits = 1)) |>
  group_by(cohort_id) |>
  fill(resp_compliance_baseline, .direction = 'updown') |>
  ungroup() 
```


```{r Create Summary Vent Table}
#First 'Long Form' Table
#Select Wanted Variables and Format
vent24_outcomes <- vent24_outcomes |>
  select(cohort_id, recorded_time, mode_category:ecmo_type, eligible_mode:resp_compliance_baseline, -temp) |>
  relocate(encounter_block, study_vent_start:lpv_end_24hours, .after = cohort_id)

#Create a Summary Table
vent24_outcome_summary <- vent24_outcomes |>
  #Get Vent Modes (For Description)
  mutate(mode_miss=fifelse(is.na(mode_category),1, 0)) |>
  group_by(cohort_id, mode_miss, eligible_tv) |>
  mutate(first_vent_mode=fcase(
    mode_miss==0 & row_number()==1 & eligible_tv==1, mode_name
  )) |>
  group_by(cohort_id) |>
  fill(first_vent_mode, .direction = 'updown') |>
  #Most Common Mode
  group_by(cohort_id, mode_name, mode_miss, eligible_tv) |>
  mutate(temp=fcase(
    mode_miss==0 & eligible_tv==1, n())) |>
  group_by(cohort_id) |>
  mutate(most_mode_tv_eligible=fcase(
    temp==max(temp, na.rm=TRUE), mode_name
  )) |>
  fill(most_mode_tv_eligible, .direction = c('downup')) |>
  ungroup()

#Select Wanted Variables and Select First Row
vent24_outcome_summary <- vent24_outcome_summary |>
  group_by(cohort_id, encounter_block) |>
  filter(row_number()==1) |>
  select(cohort_id, encounter_block, n_vent_observations, n_lpv_measures, n_tv_eligible, n_pplat_eligible, total_time_observed24, 
         time_lpv_eligible24, time_tv_eligible24, hours_at_lpv, percent_lpv, hours_at_lpv_restrict, percent_lpv_restrict, hours_tv_gt8, 
         percent_tv_gt8, hours_tv_gt7, percent_tv_gt7, hours_tv_gt6.5, percent_tv_gt6.5, hours_tv_gt6, percent_tv_gt6,
         tv_weighted24:resp_compliance_baseline, first_vent_mode, most_mode_tv_eligible) |>
  ungroup()

#Round the Time and Percent Variables
vars <- names(vent24_outcome_summary)[7:21]
vent24_outcome_summary <- vent24_outcome_summary |>
  mutate(across(
    all_of(vars),
    ~ round(., 2)
  ))

```


```{r Prepare Study ICU Names so Unit Scores Can be Merged With The Clinical Data}
PercTable(table(lpv_analytic$icu_vent), rfrq = '001')
PercTable(table(survey_unit_level$primary_icu.factor), rfrq = '001')

#Create Mapping of ICU Names
survey_unit_level <- survey_unit_level |>
  mutate(study_icu=fcase(
    primary_icu.factor=='JHH CCU, 5W', 'JHH ZAYED 5W',
    primary_icu.factor =='JHH CVSICU, 5E', 'JHH ZAYED 5E',
    primary_icu.factor =='JHH MICU, 10E', 'JHH ZAYED 10E',
    primary_icu.factor =='JHH NCCU, 3W', 'JHH ZAYED 3W',
    primary_icu.factor=='JHH WICU, Weinberg 3A', 'JHH WEINBERG 3A',
    primary_icu.factor=='JHH ONC ICU, Weinberg 5C', 'JHH WEINBERG 5C',
    primary_icu.factor=='JHH SICU, 9E', 'JHH ZAYED 9E',
    primary_icu.factor=='Sibley Memorial ICU', 'SMH ICU',
    primary_icu.factor=='Suburban 3100 ICU', 'SH 3100 INTENSIVE CARE',
    primary_icu.factor=='Suburban 3400 ICU', 'SH 3400 INTENSIVE CARE'
  )) |>
  mutate(study_icu=fifelse(is.na(study_icu), primary_icu.factor, study_icu)) |>
  select(-primary_icu.factor)

#Prepare LPV Analytic For Merging
lpv_analytic <- lpv_analytic |>
  mutate(study_icu=fcase(
    icu_vent=='SMH 2B ICU' | icu_vent=='ZDEAC SMH ICU', 'SMH ICU'
  )) |>
  mutate(study_icu=fifelse(
    is.na(study_icu), icu_vent, study_icu
  )) |>
  relocate(study_icu, .before = icu_vent)

#Merge in Unit Level Data to LPV Analytic
lpv_analytic <- lpv_analytic |>
  left_join(survey_unit_level)
```


```{r Create CCSR Categories for Analysis}
#Create List of Distinct CCSR Diagnosis Categories
ccsr_categories <- lpv_analytic |>
  select(ccsr_body_system, ccsr_category) |>
  group_by(ccsr_category) |>
  mutate(n_category=n()) |>
  ungroup() |>
  distinct() |>
  arrange(-n_category)

ccsr_system <- lpv_analytic |>
  select(ccsr_body_system) |>
  group_by(ccsr_body_system) |>
  mutate(n_category=n()) |>
  ungroup() |>
  distinct() |>
  arrange(-n_category) |>
  #Create Categorical Variable of Top 5 Systems
  mutate(ccsr_system_cat=fcase(
    ccsr_body_system=='Diseases of the circulatory system', 'Circulatory',
    ccsr_body_system=='Diseases of the respiratory system', 'Respiratory',
    ccsr_body_system %in% c('Injury, poisoning and certain other consequences of external causes',
                            'External causes of morbidity'), 'Trauma or Poisoning',
    ccsr_body_system=='Diseases of the digestive system', 'Gastrointestinal',
    ccsr_body_system=='Diseases of the nervous system', 'Neurologic',
    ccsr_body_system=='Certain infectious and parasitic diseases', 'Infectious',
    default = 'Other'
  )) |>
  select(-n_category)

#Create Categorical Variable of Top 2 Diagnoses in Each Body System
ccsr_categories <- ccsr_categories |>
  mutate(dx_category=fcase(
    ccsr_category=='Septicemia', 'Sepsis',
    ccsr_category=='Coronavirus disease – 2019 (COVID-19)', 'COVID-19',
    ccsr_category=='Cardiac arrest and ventricular fibrillation', 'Cardiac Arrest',
    ccsr_category %in% 
      c('Acute hemorrhagic cerebrovascular disease', 'Cerebral infarction',
        'Nervous system signs and symptoms', 'Epilepsy; convulsions',	'Coma; stupor; and brain damage',
        'Encephalitis'), 'Stroke, Seizure, Altered',
    ccsr_category %in% 
    c('Respiratory failure; insufficiency; arrest', 'Pneumonia (except that caused by tuberculosis)',
      'Chronic obstructive pulmonary disease and bronchiectasis', 'Other specified and unspecified lower respiratory disease',
      'Pleurisy, pleural effusion and pulmonary collapse',
      'Aspiration pneumonitis',
      'Asthma',
      'Influenza',
      'Acute bronchitis',
      'Other specified upper respiratory infections', 
      'Respiratory signs and symptoms'), 'Respiratory Failure and Pneumonia',
    ccsr_body_system=='Neoplasms', 'Cancer Related',
    ccsr_body_system=='External causes of morbidity', 'Trauma',
    ccsr_category=='Acute and unspecified renal failure', 'Acute Renal Failure',
    default = 'Other'
  ))

lpv_analytic <- lpv_analytic |>
  left_join(ccsr_categories) |>
  left_join(ccsr_system) |>
  relocate(dx_category, .after = ccsr_category) |>
  select(-n_category)
PercTable(table(lpv_analytic$dx_category), rfrq = '001')
```
```{r Add in MV Volume Per Month}
#Load MV Volume Data
mv_volume <- open_dataset(paste0(project_path, '/data/volume_of_mv.parquet')) |>
  collect()

lpv_analytic <- lpv_analytic |>
  mutate(year=year(lpv_time_zero)) |>
  mutate(month=month(lpv_time_zero)) |>
  mutate(study_month=fcase(
    year=='2022', month-6,
    year=='2023', month+6,
    year=='2024', month+18
  )) |>
  left_join(mv_volume) |>
  relocate(year, month, study_month, mv_volume_month, .after = lpv_time_zero)

```

```{r Create Study SF and PF Ratio}
#Will Use Pre Vent SF Ratios and IF Missing Post Vent SF Ratios

#For Defining PFs, can Use Following Hierarchy
#Define the P/F Ratio to Use for Adjustment Variable
#Hierarchy is: 1) Pre-vent Pao2/FiO2, 2) if that is missing the imputed pre-vent Pao2/FiO2, 3)if that is missing the Post vent PF, 4) if that is missing the imputed post vent PF
#At Each Step Describe Missingness
pf_process <- lpv_analytic |>
  mutate(pf_ratio=fcase(
    !is.na(pre_vent_pf) & pre_vent_pf<=300, pre_vent_pf
  )) 
n_pfs <- sum(!is.na(pf_process$pf_ratio))
cat('There were', n_pfs, 'qualifying PaO2/FiO2 ratios from 24 hour prior to vent \n')

#Calculate Imputed PaO2, note the Spo2 and sf_fio2 used pre vent variables if avaiable and otherwise post vent
sp2_pao2_imputation <- function(x) {
  a <- x %>%
    mutate(S=fifelse(spo2<97, spo2/100, NaN)) %>%
    mutate(a=
             11700/(1/S-1)
    ) %>%
    mutate(b =
             ((50^3+a^2)^0.5)
             ) %>%
    mutate(pao2_imputed = 
             (b+a)^0.33 - (b-a)^0.33
    ) %>%
    select(-a, -b, -S) |>
    mutate(pf_imputed=pao2_imputed/(sf_fio2/100)) |>
    relocate(pao2_imputed, pf_imputed, .after = pf_ratio)
}
pf_process <- sp2_pao2_imputation(pf_process) |>
  mutate(sf_ratio=fifelse(!is.na(pre_vent_sf), pre_vent_sf, post_vent_sf)) |>
  mutate(pf_ratio=fifelse(is.na(pf_ratio) & !is.na(pre_vent_sf), pf_imputed, pf_ratio)) #Only Uses Imputed Pre-Vent Variables

#Keep Track of What Was Imputed
cat('There were', sum(!is.na(pf_process$pf_ratio))-n_pfs, 'PaO2/FiO2 ratios using imputed data from 24 hour prior to vent \n')
cat('There were still', sum(is.na(pf_process$pf_ratio)), 'missing Pao2/FiO2 ratios \n')
n_pfs <- sum(!is.na(pf_process$pf_ratio))

pf_process <- sp2_pao2_imputation(pf_process) |>
  mutate(pf_ratio=fifelse(is.na(pf_ratio) & post_vent_pf<=300, post_vent_pf, pf_ratio)) 

cat('There were', sum(!is.na(pf_process$pf_ratio))-n_pfs, 'PaO2/FiO2 ratios directly measured in the first 12 hours after vent \n')
cat('There were still', sum(is.na(pf_process$pf_ratio)), 'missing Pao2/FiO2 ratios \n')
n_pfs <- sum(!is.na(pf_process$pf_ratio))

pf_process <- pf_process |>
  mutate(pf_ratio=fifelse(is.na(pf_ratio), pf_imputed, pf_ratio)) #Finally Takes Post Vent Imputed PF
cat('There were', sum(!is.na(pf_process$pf_ratio))-n_pfs, 'PaO2/FiO2 ratios imputed from SpO2/Fio2 in the first 12 hours after vent \n')
cat('There were still', sum(is.na(pf_process$pf_ratio)), 'missing Pao2/FiO2 ratios \n')

#Finally Describe the Pre and Post Use
pf_process <- pf_process |>
  mutate(pre_qualify=fifelse(!is.na(pre_vent_sf) | (!is.na(pre_vent_pf) & pre_vent_pf<=300), "Pre-Vent Data", "Post-Vent Data"))

cat('Table Showing Who Qualified (For Oxygenation Puposes) Using 24 hour data prior to ventilator or up to 12 hours after ventilator')
PercTable(table(pf_process$pre_qualify))

pf_process <- pf_process |>
  select(cohort_id, encounter_block, pf_ratio, sf_ratio, pre_qualify)

#Merge into LPV Analytic
lpv_analytic <- lpv_analytic |>
  left_join(pf_process) |>
  relocate(pf_ratio:pre_qualify, .after = mv_volume_month)
```

```{r Process SOFA Scores}
sofa_process <- lpv_analytic |> 
  select(cohort_id, encounter_block, max_cns_sofa:max_renal_sofa, lpv_time_zero) |>
  relocate(lpv_time_zero, .after = encounter_block) |>
  mutate(max_nr_sofa =
           rowSums(across(starts_with("max_cns_sofa"):starts_with("max_renal_sofa"))))
                   
p_missing <- unlist(lapply(sofa_process[4:9], function(x)
  sum(is.na(x))))/nrow(sofa_process)
n_missing <- unlist(lapply(sofa_process[4:9], 
                           function(x) sum(is.na(x))))
missing <- data.frame("N Missing" = n_missing, 
                 "Percent Missing" = p_missing)
print('Missinginess in SOFA Scores [Before Liver Imputation]')
missing

#Will Use Values from the 48 Hours Prior to 48 hours after LPV Time Zero for the 91 patients Missing SoFA scores
temp <- sofa_process |> filter(is.na(max_nr_sofa)) |>
  select(cohort_id, lpv_time_zero, max_coag_sofa, max_liver_sofa, max_cns_sofa)
missing_sofa <- open_dataset(paste0(project_path, '/data/lpv_master_sofa.parquet')) |> 
  filter(cohort_id %in% temp$cohort_id) |> 
  #Only Missing Values are In Bilirubin and Platelets, Filter to Those Labs Only, ONLY when A Patient Is Missing That Vlaue
  filter(lab_name=='bilirubin_total' | lab_name=='platelet_count' | !is.na(cns_sofa)) |>
  left_join(temp) |>
  filter((!is.na(coag_sofa) & is.na(max_coag_sofa)) | (!is.na(liver_sofa) & is.na(max_liver_sofa)) |
           (!is.na(cns_sofa) & is.na(max_cns_sofa))) |>
  collect() |>
  mutate(time_to_lpv=as.duration(recorded_time-lpv_time_zero)) |>
  filter(time_to_lpv>=ddays(-2) & time_to_lpv<=ddays(2)) |>
  mutate(liver=fifelse(is.na(liver_sofa), 0, 1)) |> #INdicates whether platelet or liver
  mutate(before_vent=fifelse(time_to_lpv<=ddays(0), 1, 0)) |>
  group_by(cohort_id, before_vent) |>
  mutate(templiver_sofa_max=max(liver_sofa, na.rm = TRUE)) |>
  mutate(tempcoag_sofa_max=max(coag_sofa, na.rm = TRUE)) |>
  mutate(tempcns_sofa_max=max(cns_sofa, na.rm = TRUE)) |>
  group_by(cohort_id) |>
  #Use Max Value Before Vent if Available
  mutate(liver_sofa_max=fifelse(before_vent==1, templiver_sofa_max, NaN)) |>
  mutate(coag_sofa_max=fifelse(before_vent==1, tempcoag_sofa_max, NaN)) |>
  mutate(cns_sofa_max=fifelse(before_vent==1, tempcns_sofa_max, NaN)) |>
  fill(liver_sofa_max, coag_sofa_max, cns_sofa_max, .direction = 'updown') |>
  #Can Use Max Value After Vent if Not Available
  mutate(liver_sofa_max=fifelse(is.na(liver_sofa_max), templiver_sofa_max, liver_sofa_max)) |>
  mutate(coag_sofa_max=fifelse(is.na(coag_sofa_max), tempcoag_sofa_max, coag_sofa_max)) |>
  mutate(cns_sofa_max=fifelse(is.na(cns_sofa_max), tempcns_sofa_max, cns_sofa_max)) |>
  fill(liver_sofa_max, coag_sofa_max, cns_sofa_max, .direction = 'updown') |>
  filter(row_number()==1) |>
  select(cohort_id, liver_sofa_max, coag_sofa_max, cns_sofa_max)

#Join Back with SOFA Scores 
sofa_process <- sofa_process |>
  left_join(missing_sofa) |>
  mutate(max_liver_sofa=fifelse(is.na(max_liver_sofa), liver_sofa_max, max_liver_sofa)) |>
  mutate(max_coag_sofa=fifelse(is.na(max_coag_sofa), coag_sofa_max, max_coag_sofa)) |>
  mutate(max_cns_sofa=fifelse(is.na(max_cns_sofa), cns_sofa_max, max_cns_sofa)) |>
  select(-liver_sofa_max, -coag_sofa_max, -cns_sofa_max) |>
  mutate(max_nr_sofa =
           rowSums(across(starts_with("max_cns_sofa"):starts_with("max_renal_sofa"))))
#Describe Missing
p_missing <- unlist(lapply(sofa_process[4:9], function(x)
  sum(is.na(x))))/nrow(sofa_process)
n_missing <- unlist(lapply(sofa_process[4:9], 
                           function(x) sum(is.na(x))))
missing <- data.frame("N Missing" = n_missing, 
                 "Percent Missing" = p_missing)
print('Missinginess in SOFA Scores [After Using 48 Before Window > 48 Hour After Window]')
missing

#Impute Missing Liver SOFA as Normal
sofa_process <- sofa_process |> 
  mutate(max_liver_sofa = fifelse(is.na(max_liver_sofa), 0, max_liver_sofa)) |>
  mutate(max_nr_sofa =
           rowSums(across(starts_with("max_cns_sofa"):starts_with("max_renal_sofa"))))

p_missing <- unlist(lapply(sofa_process[4:9], function(x)
  sum(is.na(x))))/nrow(sofa_process)
n_missing <- unlist(lapply(sofa_process[4:9], 
                           function(x) sum(is.na(x))))
missing <- data.frame("N Missing" = n_missing, 
                 "Percent Missing" = p_missing)
print('Missinginess in SOFA Scores [After Liver Imputation]')
missing

#Merge Back with LPV analytic
sofa_process <- sofa_process |> select(cohort_id, max_cns_sofa:max_nr_sofa)
lpv_analytic <- lpv_analytic |>
  select(-max_cns_sofa, -max_cv_sofa, -max_liver_sofa, -max_coag_sofa, -max_renal_sofa) |>
  left_join(sofa_process) |>
  relocate(max_cns_sofa:max_nr_sofa, .after=weight_kg) |>
  mutate(max_nr_sofa=fifelse(is.finite(max_nr_sofa), max_nr_sofa, NaN))
```


```{r Finally Add Vent 24 Outcomes to LPV Analytic File}
lpv_analytic <- lpv_analytic |>
  left_join(vent24_outcome_summary)
```

```{r Repeat Ventilator Outcomes for Vent Day 2}
#Primary: % Eligible Time with TV <= 6.5 cc/kg-IBW and Pplat <=30 Between Hours 24 and 48
#Need to Calculate Time At LPV and Denominator of Eligible Time
#Additionally Calculate Components (time-weighted VT per kg IBW and time-weighted PPlat)

#Track Numbers of Observations for Description in the Paper
cat("\n In the analytic cohort there were,", dim(vent72_analytic)[1], "observed ventilator settings in the first 72 hours")

vent24_48_outcomes <- vent72_analytic |> collect() |>
  #Create an End Time for LPV Assessment at 48 Hours (the earliest of ventilator end or 48 hours after lpv time zero)
  mutate(time_48hours=lpv_time_zero+dhours(48)) |>
  mutate(day2_start=lpv_time_zero+dhours(24)) |>
  mutate(lpv_end_48hours=fifelse(
      time_48hours<=study_vent_end, time_48hours, study_vent_end)) |>
  mutate(lpv_end_48hours=fifelse(
    study_vent_end<day2_start, day2_start, lpv_end_48hours)) |>
  select(-time_48hours) |>
  filter(recorded_time>day2_start & recorded_time<=lpv_end_48hours)
  
vent24_48_outcomes <- vent24_48_outcomes |>
  #How Many Ventilator Measurements on Day 2
  group_by(cohort_id) |>
  mutate(n_vent_observations=n()) |>
  #For LPV Assessment Need to Use Paired Measurements of TV and PBW
  mutate(eligible_lpv=fifelse(
    eligible_tv==1 & (!is.na(pplat_lpv) & !is.na(tv_pbw)), 1, 0
  )) |>
  group_by(cohort_id, eligible_lpv) |>
  mutate(n_lpv_measures=fifelse(
    eligible_lpv==1, n(), NaN
  )) |>
  group_by(cohort_id, eligible_tv) |>
  mutate(n_tv_eligible=fifelse(
    !is.na(tv_lpv) & eligible_tv==1, n(), NaN
  )) |>
  mutate(n_pplat_eligible=fifelse(
    !is.na(pplat_lpv) & eligible_tv==1, n(), NaN
  )) |>
  group_by(cohort_id) |>
  fill(n_lpv_measures, n_tv_eligible, n_pplat_eligible, .direction = 'updown') |>
  ungroup() |>
  mutate(n_tv_eligible=fifelse(
    is.na(n_tv_eligible), 0, n_tv_eligible
  )) |>
  mutate(n_pplat_eligible=fifelse(
    is.na(n_pplat_eligible), 0, n_pplat_eligible
  ))

#For the Patients With no Eligible PPlat Can Use Ppeak 
cat('\n Patients with no Eligible PPlat on Day 2:\n')
cat(length(unique(vent24_48_outcomes$cohort_id[vent24_48_outcomes$n_pplat_eligible==0])), 'patients \n')
cat('\n Patients with no Eligible LPV Measures in 1st 24 hours:\n')
cat(length(unique(vent24_48_outcomes$cohort_id[is.na(vent24_48_outcomes$n_lpv_measures)])), 'patients \n')

#Use Ppeak
vent24_48_outcomes <- vent24_48_outcomes |>
  #Keep Track of the Pplats in Which We Use PPeak Imputation
  mutate(pplat_imp_ppeak=fcase(
    eligible_tv==1 & is.na(n_lpv_measures), peak_inspiratory_pressure_obs
  )) |>
  #IF Pplat is Missing and It Was Otherwise an Eligible Measure Use the Ppeak Imputation
  mutate(pplat_lpv=fifelse(
    !is.na(pplat_imp_ppeak) & is.na(pplat_lpv), pplat_imp_ppeak, pplat_lpv
  )) 

#Redefine Eligible LPV and Recount the Numbers
vent24_48_outcomes <- vent24_48_outcomes |>
  mutate(eligible_lpv=fifelse(
    eligible_tv==1 & (!is.na(pplat_lpv) & !is.na(tv_pbw)), 1, 0
  )) |>
  group_by(cohort_id, eligible_lpv) |>
  mutate(n_lpv_measures=fifelse(
    eligible_lpv==1, n(), NaN
  )) |>
  group_by(cohort_id) |>
  fill(n_lpv_measures, .direction = 'updown') |>
  ungroup() |>
  mutate(n_lpv_measures=fifelse(
    is.na(n_lpv_measures), 0, n_lpv_measures
  ))

#After Imputation Counts
cat('\n In Patients with no Eligible PPlat on Day 2, We Used PPeak and After That Imputation There were:\n')
cat(length(unique(vent24_48_outcomes$cohort_id[vent24_48_outcomes$n_lpv_measures>=1])), 'patients with LPV measures for Analysis \n')
cat('\n Patients with no Eligible LPV Measures on Day 2:\n',length(unique(vent24_48_outcomes$cohort_id[vent24_48_outcomes$n_lpv_measures==0])))

#Comparison of 'Imputed PPlat' and 'Recorded Pplat'
cat('\n Recorded PPlat During Eligible Measurements \n')
summary(vent24_48_outcomes$pplat_lpv[(is.na(vent24_48_outcomes$pplat_imp_ppeak) & vent24_48_outcomes$eligible_lpv==1)])
cat('\n Ppeak Substitutin During PPlat During Eligible Measurements \n')
summary(vent24_48_outcomes$pplat_imp_ppeak)
```

```{r Create LPV Measures for Day 2 (Hours 24-48)}
#Calculate Time Spent With Each Measurement
vent24_48_outcomes <- vent24_48_outcomes |>
  #Calculate Total Time Observed on Day 2 Per Patient
  group_by(cohort_id) |>
  mutate(time_at_measure=fcase(
    row_number()!=n(), as.duration(lead(recorded_time)-recorded_time)/dhours(1), #Time to Next Observation
    row_number()==n(), as.duration(lpv_end_48hours-recorded_time)/dhours(1), #If last observation Time to End of 48 Hour Period
    recorded_time==study_vent_end, 0 #If the Last Observation is the Time of Vent End Will Call it 0 Hours *Which means exclude
  )) |>
  mutate(total_time_observed48=sum(time_at_measure)) |>
  #Carry Forward PEEP
  mutate(peep_extrapolated=peep_set) |>
  fill(peep_extrapolated, .direction= 'downup') |>
  #Now Total LPV Eligible Time
  mutate(eligible_lpv_time=fcase(
    eligible_lpv==1 | eligible_tv==0, 1, #To calculate time from paired TV/PPlat measurements to next. IF measure not eligible for TV measurement (not on controlled mode) time stops until next measurement is eligible
    eligible_lpv==0 & eligible_tv==1, 2
  )) |>
  group_by(cohort_id, eligible_lpv_time) |>
  mutate(time_at_lpv_measure=fcase(
    row_number()!=n(), as.duration(lead(recorded_time)-recorded_time)/dhours(1), #Time to Next Observation
    row_number()==n(), as.duration(lpv_end_48hours-recorded_time)/dhours(1) #If last observation Time to End of 48 Hour Period
  )) |>
  #Now Sum For Each Patient
  group_by(cohort_id, eligible_lpv) |>
  mutate(time_lpv_eligible48=fcase(
    eligible_lpv_time==1 & eligible_tv==1, sum(time_at_lpv_measure),
    default = NaN
  )) |>
  #Now Can Sum Time at TV
  group_by(cohort_id, eligible_tv) |>
  mutate(time_tv_eligible48=fcase(
    eligible_tv==1, sum(time_at_measure)
  )) |>
  #Total TV Eligible Time (NOTE: For Pplat Can Use Total LPV Time)
  group_by(cohort_id) |>
  mutate(time_tv_eligible48=fifelse(time_tv_eligible48==0, NaN, time_tv_eligible48)) |> #For Those with No Tidal Volume Time to Contribute
  mutate(time_lpv_eligible48=fifelse(time_lpv_eligible48==0, NaN, time_lpv_eligible48)) |>
  fill(time_tv_eligible48, time_lpv_eligible48, .direction='updown') |>
  ungroup()

cat('\n Time (hours) per Patient Available for LPV Analysis on Day 2[This will be the Same for Driving Pressure and PEEP \n')
summary(vent24_48_outcomes$time_lpv_eligible48[vent24_48_outcomes$n_lpv_measures>0])
hist(vent24_48_outcomes$time_lpv_eligible48[vent24_48_outcomes$n_lpv_measures>0])

cat('\n Time (hours) per Patient Available for TV Analysis on Day 2 \n')
summary(vent24_48_outcomes$time_tv_eligible48[vent24_48_outcomes$n_tv_eligible>0])
hist(vent24_48_outcomes$time_tv_eligible48[vent24_48_outcomes$n_tv_eligible>0])

#Pimary Outcome: Percentage of Time At LPV
vent24_48_outcomes <- vent24_48_outcomes |>
  mutate(met_lpv=fcase(
    eligible_lpv==1 & tv_pbw<=6.5 & pplat_lpv<=30, 1,
    eligible_lpv==1 & (tv_pbw>6.5 | pplat_lpv>30), 0
  )) |>
  #If patient Didn't Meet LPV THis Won't Be Included in Numerator so Numerator is Time LPV Met and Denominator is Total Eligible Time
  group_by(cohort_id) |>
  mutate(hours_at_lpv_day2=sum((met_lpv*time_at_lpv_measure), na.rm=TRUE)) |>
  mutate(percent_lpv_day2=
           hours_at_lpv_day2/time_lpv_eligible48) |>
  #Repeat With Restricted Definition
  mutate(met_lpv_restrict=fcase(
    eligible_lpv==1 & tv_pbw<=6.5 & pplat_lpv<=30, 1,
    eligible_lpv==1 & tv_pbw<6.0 & pplat_lpv>30, 1,
    eligible_lpv==1 & (tv_pbw>6.5 | (tv_pbw>=6.0 & pplat_lpv>30)), 0
  )) |>
  ungroup()

#Secondary Outcome Hours (%) of Time Spent with TV >8.0 in 1st 24 Hours, TV > 7 on Day 2, TV > 6.5 on Day 2
#Now Create Time-weighted TV, PEEP and Driving Pressure Measures
vent24_48_outcomes <- vent24_48_outcomes |>
  mutate(tv_gt8_day2=fcase(
    eligible_tv==1 & tv_pbw>8, 1,
    eligible_tv==1 & tv_pbw<=8, 0
  )) |>
  group_by(cohort_id) |>
  mutate(hours_tv_gt8_day2=sum((tv_gt8_day2*time_at_measure), na.rm=TRUE)) |>
  mutate(percent_tv_gt8_day2=
           hours_tv_gt8_day2/time_tv_eligible48) |>
  ungroup()
    
#Now Time Weighted TV, Pplat, PEEP and Driving Pressure; Baseline Respiratory Compliance
vent24_48_outcomes <- vent24_48_outcomes |>
  #Start with Tidal Volume
  group_by(cohort_id) |>
  #Calculate Weighted Tidal Volume = TV* (time observed/time eligible)
  mutate(temp=fcase(
    eligible_tv==1, tv_pbw*(time_at_measure/time_tv_eligible48)
  )) |>
  #Sum the Non Missing
  mutate(tv_weighted_day2=sum(temp, na.rm=TRUE)) |>
  mutate(tv_weighted_day2=round(tv_weighted_day2, digits = 1)) |>
  #Repeat For Pplat (can use eligible_lpv for this Measure)
  mutate(temp=fcase(
    eligible_lpv==1, pplat_lpv*(time_at_lpv_measure/time_lpv_eligible48)
  )) |>
  #As Along As All Values ARe Not Missing Can Do Sum
  mutate(pplat_weighted_day2=fcase(
    sum(is.na(temp))!=n(), sum(temp, na.rm=TRUE))) |>
  mutate(pplat_weighted_day2=round(pplat_weighted_day2, digits = 0)) |>
  #Repeat For PEEP (can use eligible_tv for this Measure)
  mutate(temp=fcase(
    eligible_tv==1, peep_set*(time_at_measure/time_tv_eligible48)
  )) |>
  mutate(peep_weighted_day2=fcase(
    sum(is.na(temp))!=n(), sum(temp, na.rm=TRUE))) |>
  mutate(peep_weighted_day2=round(peep_weighted_day2, digits = 0)) |>
  #Repeat For Driving Pressure (can use eligible_lpv for this Measure)
  #Filter Out Physiologically Implausible Values (Driving Pressure > 2 allowed)
  mutate(temp_dp=fifelse(
    eligible_lpv==1 & (pplat_lpv-peep_extrapolated>2), 1, 0
  )) |>
  #Filter Out the Time That the Implausible Measures Take
  mutate(temp_time=fcase(
    temp_dp==1, time_at_lpv_measure
  )) |>
  mutate(temp_time=fcase(
    sum(is.na(temp_time))!=n(), sum(temp_time, na.rm = TRUE))) |>
  mutate(temp=fcase(
    temp_dp==1, (pplat_lpv-peep_extrapolated) *(time_at_lpv_measure/temp_time)
  )) |>
  ungroup()

cat('\n Of ventilator observations on Day 2,\n ', dim(vent24_48_outcomes[vent24_48_outcomes$eligible_lpv == 1, ])[1], 'met criteria for LPV analysis,\n')
cat(' ', dim(vent24_48_outcomes[vent24_48_outcomes$eligible_tv == 1, ])[1], 'met criteria for TV analysis on Day 2, and \n')
cat(' ', dim(vent24_48_outcomes[(vent24_48_outcomes$eligible_tv == 1 & !is.na(vent24_48_outcomes$pplat_lpv)), ])[1], 'met criteria for Pplat analysis in this study. \n')

cat('\n Summary Statistics for Number of Eligible Measures Per Participant')
cat('\n Tidal Volume Measures Per Participant (Day2) \n')
summary(vent24_48_outcomes$n_tv_eligible)
cat('\n Number of Pplat Measures Per Participant (Day2) \n')
summary(vent24_48_outcomes$n_pplat_eligible)
cat('\n Number of LPV Measures Per Participant (Day2) \n')
summary(vent24_48_outcomes$n_lpv_measures)

cat('\n Patients Included in the Day2 LPV Analysis:\n')
cat(length(unique(vent24_48_outcomes$cohort_id[vent24_48_outcomes$n_lpv_measures>=1])), 'patients \n')

#Keep Track of Modes While TV Eligible
cat('\n Ventilator Modes During TV Eligible Periods on Day 2:\n')
PercTable(table(vent24_48_outcomes$mode_name[vent24_outcomes$eligible_tv==1]), rfrq = '001')

cat('\n Ventilator Modes During Any Period On Day 2\n')
PercTable(table(vent24_48_outcomes$mode_name), rfrq = '001')
```

```{r Day 2 Vent Summary}
#First 'Long Form' Table
#Select Wanted Variables and Format
vent24_48_outcomes <- vent24_48_outcomes |>
  select(cohort_id, recorded_time, mode_category:ecmo_type, eligible_mode:peep_weighted_day2, -temp) |>
  relocate(encounter_block, study_vent_start:lpv_end_48hours, .after = cohort_id)

#Create a Summary Table
vent24_48_outcome_summary <- vent24_48_outcomes |>
  #Get Vent Modes (For Description)
  mutate(mode_miss=fifelse(is.na(mode_category),1, 0)) |>
  group_by(cohort_id, mode_miss, eligible_tv) |>
  mutate(first_vent_mode=fcase(
    mode_miss==0 & row_number()==1 & eligible_tv==1, mode_name
  )) |>
  group_by(cohort_id) |>
  fill(first_vent_mode, .direction = 'updown') |>
  #Most Common Mode
  group_by(cohort_id, mode_name, mode_miss, eligible_tv) |>
  mutate(temp=fcase(
    mode_miss==0 & eligible_tv==1, n())) |>
  group_by(cohort_id) |>
  mutate(most_mode_tv_eligible=fcase(
    temp==max(temp, na.rm=TRUE), mode_name
  )) |>
  fill(most_mode_tv_eligible, .direction = c('downup')) |>
  ungroup()

#Select Wanted Variables and Select First Row
vent24_48_outcome_summary <- vent24_48_outcome_summary |>
  group_by(cohort_id, encounter_block) |>
  filter(row_number()==1) |>
  select(cohort_id, encounter_block, n_vent_observations, n_lpv_measures, n_tv_eligible, n_pplat_eligible, total_time_observed48, 
         time_lpv_eligible48, time_tv_eligible48, hours_at_lpv_day2, percent_lpv_day2, hours_tv_gt8_day2, 
         percent_tv_gt8_day2, tv_weighted_day2:peep_weighted_day2) |>
  mutate(hours_at_lpv_day2=fifelse(is.na(time_lpv_eligible48), NaN, hours_at_lpv_day2)) |>
  mutate(hours_tv_gt8_day2=fifelse(is.na(time_tv_eligible48), NaN, hours_tv_gt8_day2)) |>
  mutate(tv_weighted_day2=fifelse(is.na(time_tv_eligible48), NaN, tv_weighted_day2)) |>
  ungroup()

#Round the Time and Percent Variables
vars <- names(vent24_48_outcome_summary)[7:13]
vent24_48_outcome_summary <- vent24_48_outcome_summary |>
  mutate(across(
    all_of(vars),
    ~ round(., 2)
  ))

#Format for Merge
vent24_48_outcome_summary <- vent24_48_outcome_summary |>
  filter(total_time_observed48>0) |>
  filter(n_tv_eligible>0 | n_lpv_measures>0) |>
  rename(n_vent_observations_day2=n_vent_observations,
         n_lpv_measures_day2=n_lpv_measures, 
         n_tv_eligible_day2=n_tv_eligible,
         n_pplat_eligible_day2=n_pplat_eligible)
```


```{r Vent Day 3 Outcomes}
#Primary: % Eligible Time with TV <= 6.5 cc/kg-IBW and Pplat <=30 Between Hours 24 and 48
#Need to Calculate Time At LPV and Denominator of Eligible Time
#Additionally Calculate Components (time-weighted VT per kg IBW and time-weighted PPlat)

#Track Numbers of Observations for Description in the Paper
cat("\n In the analytic cohort there were,", dim(vent72_analytic)[1], "observed ventilator settings in the first 72 hours")

vent48_72_outcomes <- vent72_analytic |> collect() |>
  #Create an End Time for LPV Assessment at 72 Hours (the earliest of ventilator end or 72 hours after lpv time zero)
  mutate(time_72hours=lpv_time_zero+dhours(72)) |>
  mutate(day3_start=lpv_time_zero+dhours(48)) |>
  mutate(lpv_end_72hours=fifelse(
      time_72hours<=study_vent_end, time_72hours, study_vent_end)) |>
  mutate(lpv_end_72hours=fifelse(
    study_vent_end<day3_start, day3_start, lpv_end_72hours)) |>
  select(-time_72hours) |>
  filter(recorded_time>day3_start & recorded_time<=lpv_end_72hours)
  
vent48_72_outcomes <- vent48_72_outcomes |>
  #How Many Ventilator Measurements on Day 3
  group_by(cohort_id) |>
  mutate(n_vent_observations=n()) |>
  #For LPV Assessment Need to Use Paired Measurements of TV and PBW
  mutate(eligible_lpv=fifelse(
    eligible_tv==1 & (!is.na(pplat_lpv) & !is.na(tv_pbw)), 1, 0
  )) |>
  group_by(cohort_id, eligible_lpv) |>
  mutate(n_lpv_measures=fifelse(
    eligible_lpv==1, n(), NaN
  )) |>
  group_by(cohort_id, eligible_tv) |>
  mutate(n_tv_eligible=fifelse(
    !is.na(tv_lpv) & eligible_tv==1, n(), NaN
  )) |>
  mutate(n_pplat_eligible=fifelse(
    !is.na(pplat_lpv) & eligible_tv==1, n(), NaN
  )) |>
  group_by(cohort_id) |>
  fill(n_lpv_measures, n_tv_eligible, n_pplat_eligible, .direction = 'updown') |>
  ungroup() |>
  mutate(n_tv_eligible=fifelse(
    is.na(n_tv_eligible), 0, n_tv_eligible
  )) |>
  mutate(n_pplat_eligible=fifelse(
    is.na(n_pplat_eligible), 0, n_pplat_eligible
  ))

#For the Patients With no Eligible PPlat Can Use Ppeak 
cat('\n Patients with no Eligible PPlat on Day 2:\n')
cat(length(unique(vent48_72_outcomes$cohort_id[vent48_72_outcomes$n_pplat_eligible==0])), 'patients \n')
cat('\n Patients with no Eligible LPV Measures in 1st 24 hours:\n')
cat(length(unique(vent48_72_outcomes$cohort_id[is.na(vent48_72_outcomes$n_lpv_measures)])), 'patients \n')

#Use Ppeak
vent48_72_outcomes <- vent48_72_outcomes |>
  #Keep Track of the Pplats in Which We Use PPeak Imputation
  mutate(pplat_imp_ppeak=fcase(
    eligible_tv==1 & is.na(n_lpv_measures), peak_inspiratory_pressure_obs
  )) |>
  #IF Pplat is Missing and It Was Otherwise an Eligible Measure Use the Ppeak Imputation
  mutate(pplat_lpv=fifelse(
    !is.na(pplat_imp_ppeak) & is.na(pplat_lpv), pplat_imp_ppeak, pplat_lpv
  )) 

#Redefine Eligible LPV and Recount the Numbers
vent48_72_outcomes <- vent48_72_outcomes |>
  mutate(eligible_lpv=fifelse(
    eligible_tv==1 & (!is.na(pplat_lpv) & !is.na(tv_pbw)), 1, 0
  )) |>
  group_by(cohort_id, eligible_lpv) |>
  mutate(n_lpv_measures=fifelse(
    eligible_lpv==1, n(), NaN
  )) |>
  group_by(cohort_id) |>
  fill(n_lpv_measures, .direction = 'updown') |>
  ungroup() |>
  mutate(n_lpv_measures=fifelse(
    is.na(n_lpv_measures), 0, n_lpv_measures
  ))

#After Imputation Counts
cat('\n In Patients with no Eligible PPlat on Day 3, We Used PPeak and After That Imputation There were:\n')
cat(length(unique(vent48_72_outcomes$cohort_id[vent48_72_outcomes$n_lpv_measures>=1])), 'patients with LPV measures for Analysis \n')
cat('\n Patients with no Eligible LPV Measures on Day 2:\n',length(unique(vent48_72_outcomes$cohort_id[vent48_72_outcomes$n_lpv_measures==0])))

#Comparison of 'Imputed PPlat' and 'Recorded Pplat'
cat('\n Recorded PPlat During Eligible Measurements \n')
summary(vent48_72_outcomes$pplat_lpv[(is.na(vent48_72_outcomes$pplat_imp_ppeak) & vent48_72_outcomes$eligible_lpv==1)])
cat('\n Ppeak Substitutin During PPlat During Eligible Measurements \n')
summary(vent48_72_outcomes$pplat_imp_ppeak)
```
```{r Vent Day 3 OUtcomes}
#Calculate Time Spent With Each Measurement
vent48_72_outcomes <- vent48_72_outcomes |>
  #Calculate Total Time Observed on Day 3 Per Patient
  group_by(cohort_id) |>
  mutate(time_at_measure=fcase(
    row_number()!=n(), as.duration(lead(recorded_time)-recorded_time)/dhours(1), #Time to Next Observation
    row_number()==n(), as.duration(lpv_end_72hours-recorded_time)/dhours(1), #If last observation Time to End of 72 Hour Period
    recorded_time==study_vent_end, 0 #If the Last Observation is the Time of Vent End Will Call it 0 Hours *Which means exclude
  )) |>
  mutate(total_time_observed72=sum(time_at_measure)) |>
  #Carry Forward PEEP
  mutate(peep_extrapolated=peep_set) |>
  fill(peep_extrapolated, .direction= 'downup') |>
  #Now Total LPV Eligible Time
  mutate(eligible_lpv_time=fcase(
    eligible_lpv==1 | eligible_tv==0, 1, #To calculate time from paired TV/PPlat measurements to next. IF measure not eligible for TV measurement (not on controlled mode) time stops until next measurement is eligible
    eligible_lpv==0 & eligible_tv==1, 2
  )) |>
  group_by(cohort_id, eligible_lpv_time) |>
  mutate(time_at_lpv_measure=fcase(
    row_number()!=n(), as.duration(lead(recorded_time)-recorded_time)/dhours(1), #Time to Next Observation
    row_number()==n(), as.duration(lpv_end_72hours-recorded_time)/dhours(1) #If last observation Time to End of 72 Hour Period
  )) |>
  #Now Sum For Each Patient
  group_by(cohort_id, eligible_lpv) |>
  mutate(time_lpv_eligible72=fcase(
    eligible_lpv_time==1 & eligible_tv==1, sum(time_at_lpv_measure),
    default = NaN
  )) |>
  #Now Can Sum Time at TV
  group_by(cohort_id, eligible_tv) |>
  mutate(time_tv_eligible72=fcase(
    eligible_tv==1, sum(time_at_measure)
  )) |>
  #Total TV Eligible Time (NOTE: For Pplat Can Use Total LPV Time)
  group_by(cohort_id) |>
  mutate(time_tv_eligible72=fifelse(time_tv_eligible72==0, NaN, time_tv_eligible72)) |> #For Those with No Tidal Volume Time to Contribute
  mutate(time_lpv_eligible72=fifelse(time_lpv_eligible72==0, NaN, time_lpv_eligible72)) |>
  fill(time_tv_eligible72, time_lpv_eligible72, .direction='updown') |>
  ungroup()

cat('\n Time (hours) per Patient Available for LPV Analysis on Day 3[This will be the Same for Driving Pressure and PEEP \n')
summary(vent48_72_outcomes$time_lpv_eligible72[vent48_72_outcomes$n_lpv_measures>0])
hist(vent48_72_outcomes$time_lpv_eligible72[vent48_72_outcomes$n_lpv_measures>0])

cat('\n Time (hours) per Patient Available for TV Analysis on Day 3 \n')
summary(vent48_72_outcomes$time_tv_eligible72[vent48_72_outcomes$n_tv_eligible>0])
hist(vent48_72_outcomes$time_tv_eligible72[vent48_72_outcomes$n_tv_eligible>0])

#Pimary Outcome: Percentage of Time At LPV
vent48_72_outcomes <- vent48_72_outcomes |>
  mutate(met_lpv=fcase(
    eligible_lpv==1 & tv_pbw<=6.5 & pplat_lpv<=30, 1,
    eligible_lpv==1 & (tv_pbw>6.5 | pplat_lpv>30), 0
  )) |>
  #If patient Didn't Meet LPV THis Won't Be Included in Numerator so Numerator is Time LPV Met and Denominator is Total Eligible Time
  group_by(cohort_id) |>
  mutate(hours_at_lpv_day3=sum((met_lpv*time_at_lpv_measure), na.rm=TRUE)) |>
  mutate(percent_lpv_day3=
           hours_at_lpv_day3/time_lpv_eligible72) |>
  #Repeat With Restricted Definition
  mutate(met_lpv_restrict=fcase(
    eligible_lpv==1 & tv_pbw<=6.5 & pplat_lpv<=30, 1,
    eligible_lpv==1 & tv_pbw<6.0 & pplat_lpv>30, 1,
    eligible_lpv==1 & (tv_pbw>6.5 | (tv_pbw>=6.0 & pplat_lpv>30)), 0
  )) |>
  ungroup()

#Secondary Outcome Hours (%) of Time Spent with TV >8.0 in 1st 24 Hours, TV > 7 on Day 2, TV > 6.5 on Day 2
#Now Create Time-weighted TV, PEEP and Driving Pressure Measures
vent48_72_outcomes <- vent48_72_outcomes |>
  mutate(tv_gt8_day3=fcase(
    eligible_tv==1 & tv_pbw>8, 1,
    eligible_tv==1 & tv_pbw<=8, 0
  )) |>
  group_by(cohort_id) |>
  mutate(hours_tv_gt8_day3=sum((tv_gt8_day3*time_at_measure), na.rm=TRUE)) |>
  mutate(percent_tv_gt8_day3=
           hours_tv_gt8_day3/time_tv_eligible72) |>
  ungroup()
    
#Now Time Weighted TV, Pplat, PEEP and Driving Pressure; Baseline Respiratory Compliance
vent48_72_outcomes <- vent48_72_outcomes |>
  #Start with Tidal Volume
  group_by(cohort_id) |>
  #Calculate Weighted Tidal Volume = TV* (time observed/time eligible)
  mutate(temp=fcase(
    eligible_tv==1, tv_pbw*(time_at_measure/time_tv_eligible72)
  )) |>
  #Sum the Non Missing
  mutate(tv_weighted_day3=sum(temp, na.rm=TRUE)) |>
  mutate(tv_weighted_day3=round(tv_weighted_day3, digits = 1)) |>
  #Repeat For Pplat (can use eligible_lpv for this Measure)
  mutate(temp=fcase(
    eligible_lpv==1, pplat_lpv*(time_at_lpv_measure/time_lpv_eligible72)
  )) |>
  #As Along As All Values ARe Not Missing Can Do Sum
  mutate(pplat_weighted_day3=fcase(
    sum(is.na(temp))!=n(), sum(temp, na.rm=TRUE))) |>
  mutate(pplat_weighted_day3=round(pplat_weighted_day3, digits = 0)) |>
  #Repeat For PEEP (can use eligible_tv for this Measure)
  mutate(temp=fcase(
    eligible_tv==1, peep_set*(time_at_measure/time_tv_eligible72)
  )) |>
  mutate(peep_weighted_day3=fcase(
    sum(is.na(temp))!=n(), sum(temp, na.rm=TRUE))) |>
  mutate(peep_weighted_day3=round(peep_weighted_day3, digits = 0)) |>
  #Repeat For Driving Pressure (can use eligible_lpv for this Measure)
  #Filter Out Physiologically Implausible Values (Driving Pressure > 2 allowed)
  mutate(temp_dp=fifelse(
    eligible_lpv==1 & (pplat_lpv-peep_extrapolated>2), 1, 0
  )) |>
  #Filter Out the Time That the Implausible Measures Take
  mutate(temp_time=fcase(
    temp_dp==1, time_at_lpv_measure
  )) |>
  mutate(temp_time=fcase(
    sum(is.na(temp_time))!=n(), sum(temp_time, na.rm = TRUE))) |>
  mutate(temp=fcase(
    temp_dp==1, (pplat_lpv-peep_extrapolated) *(time_at_lpv_measure/temp_time)
  )) |>
  ungroup()

cat('\n Of ventilator observations on Day 2,\n ', dim(vent48_72_outcomes[vent48_72_outcomes$eligible_lpv == 1, ])[1], 'met criteria for LPV analysis,\n')
cat(' ', dim(vent48_72_outcomes[vent48_72_outcomes$eligible_tv == 1, ])[1], 'met criteria for TV analysis on Day 2, and \n')
cat(' ', dim(vent48_72_outcomes[(vent48_72_outcomes$eligible_tv == 1 & !is.na(vent48_72_outcomes$pplat_lpv)), ])[1], 'met criteria for Pplat analysis in this study. \n')

cat('\n Summary Statistics for Number of Eligible Measures Per Participant')
cat('\n Tidal Volume Measures Per Participant (day3) \n')
summary(vent48_72_outcomes$n_tv_eligible)
cat('\n Number of Pplat Measures Per Participant (day3) \n')
summary(vent48_72_outcomes$n_pplat_eligible)
cat('\n Number of LPV Measures Per Participant (day3) \n')
summary(vent48_72_outcomes$n_lpv_measures)

cat('\n Patients Included in the day3 LPV Analysis:\n')
cat(length(unique(vent48_72_outcomes$cohort_id[vent48_72_outcomes$n_lpv_measures>=1])), 'patients \n')

#Keep Track of Modes While TV Eligible
cat('\n Ventilator Modes During TV Eligible Periods on Day 2:\n')
PercTable(table(vent48_72_outcomes$mode_name[vent24_outcomes$eligible_tv==1]), rfrq = '001')

cat('\n Ventilator Modes During Any Period On Day 2\n')
PercTable(table(vent48_72_outcomes$mode_name), rfrq = '001')
```


```{r Day 3 Vent Summary}
#First 'Long Form' Table
#Select Wanted Variables and Format
vent48_72_outcomes <- vent48_72_outcomes |>
  select(cohort_id, recorded_time, mode_category:ecmo_type, eligible_mode:peep_weighted_day3, -temp) |>
  relocate(encounter_block, study_vent_start:lpv_end_72hours, .after = cohort_id)

#Create a Summary Table
vent48_72_outcome_summary <- vent48_72_outcomes |>
  #Get Vent Modes (For Description)
  mutate(mode_miss=fifelse(is.na(mode_category),1, 0)) |>
  group_by(cohort_id, mode_miss, eligible_tv) |>
  mutate(first_vent_mode=fcase(
    mode_miss==0 & row_number()==1 & eligible_tv==1, mode_name
  )) |>
  group_by(cohort_id) |>
  fill(first_vent_mode, .direction = 'updown') |>
  #Most Common Mode
  group_by(cohort_id, mode_name, mode_miss, eligible_tv) |>
  mutate(temp=fcase(
    mode_miss==0 & eligible_tv==1, n())) |>
  group_by(cohort_id) |>
  mutate(most_mode_tv_eligible=fcase(
    temp==max(temp, na.rm=TRUE), mode_name
  )) |>
  fill(most_mode_tv_eligible, .direction = c('downup')) |>
  ungroup()

#Select Wanted Variables and Select First Row
vent48_72_outcome_summary <- vent48_72_outcome_summary |>
  group_by(cohort_id, encounter_block) |>
  filter(row_number()==1) |>
  select(cohort_id, encounter_block, n_vent_observations, n_lpv_measures, n_tv_eligible, n_pplat_eligible, total_time_observed72, 
         time_lpv_eligible72, time_tv_eligible72, hours_at_lpv_day3, percent_lpv_day3, hours_tv_gt8_day3, 
         percent_tv_gt8_day3, tv_weighted_day3:peep_weighted_day3) |>
  mutate(hours_at_lpv_day3=fifelse(is.na(time_lpv_eligible72), NaN, hours_at_lpv_day3)) |>
  mutate(hours_tv_gt8_day3=fifelse(is.na(time_tv_eligible72), NaN, hours_tv_gt8_day3)) |>
  mutate(tv_weighted_day3=fifelse(is.na(time_tv_eligible72), NaN, tv_weighted_day3)) |>
  ungroup()

#Round the Time and Percent Variables
vars <- names(vent48_72_outcome_summary)[7:13]
vent48_72_outcome_summary <- vent48_72_outcome_summary |>
  mutate(across(
    all_of(vars),
    ~ round(., 2)
  ))

#Format for Merge
vent48_72_outcome_summary <- vent48_72_outcome_summary |>
  filter(total_time_observed72>0) |>
  filter(n_tv_eligible>0 | n_lpv_measures>0) |>
  rename(n_vent_observations_day3=n_vent_observations,
         n_lpv_measures_day3=n_lpv_measures, 
         n_tv_eligible_day3=n_tv_eligible,
         n_pplat_eligible_day3=n_pplat_eligible)
```


```{r Add Day 2 and 3 Summaries to LPV Analytic}
lpv_analytic <- lpv_analytic |>
  left_join(vent24_48_outcome_summary) |>
  left_join(vent48_72_outcome_summary)

```


```{r Save Analytic Files}
write_parquet(lpv_analytic, paste0(project_path, '/data/lpv_analytic.parquet'))
write_parquet(vent24_outcome_summary, paste0(project_path, '/data/vent24_outcome_summary.parquet'))
write_parquet(vent24_outcomes, paste0(project_path, '/data/vent24_outcomes_long.parquet'))
```

